Automatic Matching of Bullet Lands

### 2016-01-15 imaging-paper.Rnw
moved the CMS discussion to the proper place

All of the other steps are done on pairs of bullet lands:
\begin{enumerate}
\item {\bf Smooth the two signatures} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
\item Use cross-correlation to {\bf find the best alignment} of the two signatures: shift one of the signatures by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
\item Using a rolling average, {\bf identify peaks and valleys} for each of the signatures. We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}). Peaks and valleys constitute the \emph{striation marks} of the bullet.
\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as those areas, in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals. A joint interval is then called a match(ing stria) between the signatures, if all of the intervals are of the same type of extrema, i.e.\ they are either all peaks or all valleys. In Figure~\ref{fig:match} all matches are shown as color-filled rectangles corresponding to their type of extrema (peaks are shown in orange, and valleys in green). Non-matching intervals are left grey. \hh{XXX this definition does not coincide with the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks. Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands. The definition given in this paper, is computationally far less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land. In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). XXX}
\item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted. Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
\begin{enumerate}[label=(\roman*)]
\item Maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
\item Number of matches and non-matches,
\item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
\[
D^2 = \frac{1}{\text{\#}t}\sum_t \left[f(t) - g(t)\right]^2,
\]
\item The sum $S$ of average absolute heights of matched extrema: for each of the two matched stria, compute the average of the absolute heights of the peaks or valleys. $S$ is then defined as the sum of all these averages. 
\end{enumerate}
\end{enumerate}
%
The difference $D$ between signatures is here defined as the Euclidean distance (in $\mu m$). In the paper by \citet{ma:2004}, distance is defined as a measure relative to the first signature, which serves as a comparison reference and is therefore a unitless quantity.

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. We will assess this claim in the next section.

>>>>

All of the other steps are done on pairs of bullet lands:
\begin{enumerate}
\item {\bf Smooth the two signatures} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
\item Use cross-correlation to {\bf find the best alignment} of the two signatures: shift one of the signatures by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
\item Using a rolling average, {\bf identify peaks and valleys} for each of the signatures. We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}). Peaks and valleys constitute the \emph{striation marks} of the bullet.
\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as those areas, in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals. A joint interval is then called a match(ing stria) between the signatures, if all of the intervals are of the same type of extrema, i.e.\ they are either all peaks or all valleys. In Figure~\ref{fig:match} all matches are shown as color-filled rectangles corresponding to their type of extrema (peaks are shown in orange, and valleys in green). Non-matching intervals are left grey. 
\item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted. Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
\begin{enumerate}[label=(\roman*)]
\item Maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
\item Number of matches and non-matches,
\item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
\[
D^2 = \frac{1}{\text{\#}t}\sum_t \left[f(t) - g(t)\right]^2,
\]
\item The sum $S$ of average absolute heights of matched extrema: for each of the two matched stria, compute the average of the absolute heights of the peaks or valleys. $S$ is then defined as the sum of all these averages. 
\end{enumerate}
\end{enumerate}
%
The difference $D$ between signatures is here defined as the Euclidean distance (in $\mu m$). In the paper by \citet{ma:2004}, distance is defined as a measure relative to the first signature, which serves as a comparison reference and is therefore a unitless quantity.

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
\hh{XXX Note that the definition given in this paper does not coincide with the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks. Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands. The definition given in here, is computationally less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land. In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). XXX}
For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. \hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition. Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley? XXX} We will assess this claim in the next section.
